### 2015-12-04 paper/test.tex
writing study setup and case study

\documentclass[conference]{IEEEtran}
%\usepackage[top=1.2in, bottom=1.5in, left=0.8in, right=0.8in]{geometry}
\usepackage{balance}
\usepackage{graphicx}
\usepackage{xcolor}
\newcommand{\todo}[1]{\qquad\textcolor{red}{TODO: #1}\PackageWarning{TODO:}{#1!}}

\hyphenation{op-tical net-works semi-conduc-tor}

\makeatletter
\setlength{\@fptop}{0pt}
\makeatother

\begin{document}
\title{Togglge Smell Detection in Google Chrome}
\author{
Md Tajmilur Rahman, Dharani Kumar, Sumir Sarkar, Omar Faruk Miah\\
Concordia University,~Montreal,~QC,~Canada
}
\maketitle

\begin{abstract}
TODO
\end{abstract}

\begin{IEEEkeywords}
Software Refactoring, Software Re-Engineering, Software Design, Feature Toggles, Code Smell, Technical Debt, Mining Repository
\end{IEEEkeywords}